### 2023-10-21 docs/manuscript/abstract.tex
Extract some sections into their own files.

    We study the classic problem of inferring ancestral genomes on a given phylogeny and a given set of extant genomes, also known as \emph{small parsimony problem} (SPP). 
    More specifically, we address a scenario where genomes represent sets of linear or circular successions of oriented genomic markers orders, also known as \emph{gene orders} that evolve along the branches of the phylogeny and are altered by a general evolutionary model that supports a genome rearrangements emulated through the double-cut-and-join (DCJ) operation and by segmental insertions and deletions.

    Even under simple evolutionary models, this problem is computationally intractable,